---
title: "Reproducible Research: Peer Assessment 1"
output: html_document
keep_md: true
---

Aurthor: Zaw Htet Wai  
Source code repo:  
https://github.com/zawhtetwai/RepData_PeerAssessment1   
Reproducible output:  
https://github.com/zawhtetwai/RepData_PeerAssessment1/blob/master/PA1_template.html

## Loading and preprocessing the data  
The source data for this assignment can be downloaded from the course web site:  
Dataset: [Activity monitoring data](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip) [52K]  

Below is the code to unzip and read the data as the CSV file.  
And have a look at its summary.  

```{r, echo=TRUE}
library(plyr)
library(ggplot2)
# Read the data from zip file  
activity <- read.csv(unz("activity.zip", "activity.csv"))
# Preview the summary of the data
summary(activity)
```

## What is mean total number of steps taken per day?  
Calculate the total steps are taken each day for the analysis  
1. Plot the analysis as the histogram  

```{r, echo=TRUE}
total <- ddply(activity, .(date), summarize, steps=sum(steps))
barplot(total$steps, names.arg = total$date, main ="Number of Steps Taken Each Day\n", xlab ="Date", ylab = "Steps")
```

2. Mean and median based on the total number of steps taken per day
```{r, echo=TRUE}
mean_steps <- mean(total$steps, na.rm=TRUE)
round(mean_steps, digits = 0)
median_steps <- median(total$steps, na.rm=TRUE)
round(median_steps, digits = 0)
```


## What is the average daily activity pattern?  
1. Calculate the average number of steps taken and create the time series plot of the 5-minute interval on average across all the days in the dataset  


```{r, echo=TRUE}
daily_avg = aggregate(steps ~ interval, activity, mean)
ggplot(daily_avg, aes(x = interval, y = steps)) + geom_line() + xlab("5-minute Interval") + ylab("Steps")
```

2. Find the 5-minute interval that contains the maximum number of steps  

```{r, echo=TRUE}
daily_avg[which.max(daily_avg$steps), "interval"]
```

## Imputing missing values  
1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)  

```{r, echo=TRUE}
missing_val <- sum(is.na(activity))
print(missing_val)
```
There are total `r missing_val` missing values.  

2. Following the instructions in the assignment,  
**I propose the imputing strategy of filling these missing values with their mean of 5-minute interval on that day**.  

```{r, echo=TRUE}
impute_func <- function(x, y) {    
    if (!is.na(x)) 
        res <- c(as.numeric(x))
     else res <- as.numeric(daily_avg[daily_avg$interval == y, "steps"])
    return(res)
}
```

3. Below code create a new dataset that is equal to the original dataset but with the missing data filled in.  
And have a look at the new dataset summary.

```{r, echo=TRUE}
new_data <- activity
new_data$steps <- mapply(impute_func, new_data$steps, new_data$interval)
summary(new_data)
```

Then check to ensure there is no record with missing value (NA)  
```{r, echo=TRUE}
sum(is.na(new_data))
```

## Are there differences in activity patterns between weekdays and weekends?
